\documentclass{article}
\usepackage{amsmath}
\begin{document}

% Outline:
% 1. Introduction
% 2. Methods
%  - PK model (two-compartment)
%  - PD summaries (T>MIC)
%  - Statistical model (Bayesian NLR)
% Quality of predictions (computing efficiency, 'coverage' of credible intervals)
% Extensions
% Conclusions

\section{Introduction}
\begin{itemize}
\item Goal is to facilitate real-time herapeutic drug monitoring (TDM); translational research.
\item PK predictions often not reported with uncertainty; cite literature.
\item Need to quantify uncertianty in real time, but it's difficult.
\item incorporate objective prior information
\item We created a web app and suite of methods that accomplish this.
\end{itemize}

\section{Methods}
\subsection{Two-compartment model}
The two-compartment pharmacokinetic model is expressed as a system of two ordinary differential equations as follows, where $m_1$ and $m_2$ are the masses of drug in the central and peripheral compartments, respectively.

\begin{align}
\frac{dm_1}{dt} &= -k_{10}m_1 - k_{12}m_1 + k_{21}m_2 + k_R \nonumber \\
\frac{dm_2}{dt} &= \phantom{-k_{10}m_1} + k_{12}m_1 - k_{21}m_2 \nonumber
\end{align}

The concentration of drug in the central compartment is given by $c_1 = m_1/v_1$, where $v_1$ is the volume of the central compartment. The parameters $k_{10}$, $k_{12}$, $k_{21}$, and $k_R$ are described in Table \ref{tab:pkpars}

\begin{table}
\begin{tabular}{lll} \hline
Parameter & Units & Description \\ \hline
$k_{10}$ & h$^{-1}$ & Elimination rate from central compartment\\
$k_{12}$ & h$^{-1}$ & Distribution rate from central to peripheral compartment\\
$k_{21}$ & h$^{-1}$ & Distribution rate from peripheral to central compartment\\
$k_R$  & g$\cdot$h$^{-1}$ & Infusion rate into central compartment\\
$v_1$  & L & Volume of central compartment\\
\hline
\end{tabular}
\caption{Two-compartment model parameter units (SI) and descriptions. \label{tab:pkpars}}
\end{table}

\subsection{Bayes prediction model}
Concentration measurements are modeled using a nonlinear regression method with either additive error as follows:
\begin{displaymath}
c_{ij} = \eta_i(t_{ij}, \theta_i) + \epsilon_{ij}
\end{displaymath}
\noindent In this expression, $c_{ij}$ is the measured concentration for subject $i = 1 \ldots n$ at time $t_{ij}$ for $j = 1 \ldots m_i$, $\eta_i(t_{ij}, \theta_i)$ is the two-compartment model solution for subject $i$ at time $t_{ij}$ given parameters $\theta_i = [k_{10i}, k_{12i}, k_{21i}, v_{1i}]$, and $\epsilon_{ij}$ represents i.i.d. random error with mean zero. By specifying the random error density function, say $f(\epsilon_{ij}, \sigma)$, the subject-specific likelihood function is 
\begin{displaymath}
L_i(\theta_i, \sigma) = \prod_{j=1}^{m_i} f(c_{ij} - \eta_i(t_{ij}, \theta_i), \sigma).
\end{displaymath}
\noindent Thus, given a prior distribution $\pi_0(\theta_i)$, the subject-specific posterior is proportional to $\pi_i(\theta_i, \sigma) \propto \pi_0(\theta_i,\sigma)L_i(\theta_i,\sigma)$.

In the current context, $f$ is taken as the normal density function with mean zero and standard deviation $\sigma$, and the prior distribution is generated to satisfy the following:
\begin{align}
[\log \theta_i] &\sim N_4(\mu_0, \Sigma_0) \\
[\log \sigma] &\sim N_1(m_0, s_0)
\end{align}
\noindent where $N_4(\mu_0, \Sigma_0)$ represents the 4-variate normal distribution with mean $\mu_0$ and covariance matrix $\Sigma_0$, and $N_1(m_0, s_0) represents the univariate normal distribution with mean $m_0$ and variance $s_0$. The prior distribution for the PK parameters represents our prior knowledge about the PK heterogeneity for a particular drug in a target population. Thus, the values of the prior hyperparameters should be carefully selected for the task at hand. In the current context, and by default, the hyperparameters are specified to correspond with estimates that arose from a study of piperacillin pharmacokinetics in a hospitalized, critically ill population. The particular values are listed in an appendix. 

Due to the nonlinearity of the two-compartment model, the posterior distribution does not take a familiar form, and posterior summaries must be approximated. In particular, we sought to compute 95% credible bands for subject-specific concentration-time curves, and for a specific summary of the curve: the length of time in which the concentration exceeds a specified value. Monte Carlo techniques are often used to approximate these quantities. However, because these posterior summaries are presented in a web application in real time, we sought alternatives that were less computationally intensive and deterministic. We consider several methods as described in the following sections.

\subsubsection{Laplace approximation}
We first considered a method that makes use of two approximations. The first is a Laplace approximation to the subject-specific posterior density, and the second is a first order Taylor approximation of the target summary (i.e., the `delta method'). The Laplace approximation is given as follows
\begin{displaymath}
\pi_i(\log \theta_i, \log \sigma) \approx N([\log \hat{\theta_i},\log \hat{\sigma}], [-H]^{-1})
\end{displaymath} 
\noindent where $[\log \hat{\theta_i},\log \hat{\sigma}]$ is the posterior mode and $H_{\hat{\theta_i}}$ is the posterior Hessian with respect to $[\log \theta_i, \log \sigma]$ evaluated at the posterior mode. The second approximation makes use of the delta method, such that a posterior functional $h(\log \theta_i)$ has an approximate normal distribution. In the present context, $h(\log \theta_i)$ represents the logit of the fraction of the dosing period in which the concentration of drug exceeds a spedified value. The first-order Taylor expansion of $h(\log \theta_i)$ is 
\begin{displaymath}
h(\log \theta_i) \approx  h(\log \hat{\theta_i}) + G^T(\log \theta_i - \log \hat{\theta_i}),
\end{displaymath}
\noindent where $G$ is the gradient of $h(\log \theta_i)$ with respect to $[\log \theta_i, \log \sigma]$ evaluated at the posterior mode. Thus, given that $[\log \theta_i, \log \sigma]$ has an approximate normal distribution, the same is true for $h(\log \theta_i)$:
\begin{displaymath}
h(\log \theta_i) \sim N(h(\log \hat{\theta_i}), G^T[-H]^{-1}G). 
\end{displaymath}
\noindent An approximate $(1-\alpha)\cdot 100$\% credible interval for $h(\log \theta_i)$ is thus given by the $\alpha/2$ and $(1-\alpha)/2$ quantiles of the approximate posterior distribution for $h(\log \theta)$. This method is computationally elegant, since the posterior Hessian and posterior mode can be computed simultaneously by most optimization software routines (e.g., the R function `optim'). Indeed, this method usually has the smallest computational burden (i.e., number of likelihood evaluations) among the three methods considered here.

\subsubsection{Variational approximation}
\subsubsection{Quadrature approximation}
\subsection{Evaluation of approximation methods}
\begin{itemize}
\item used MCMC as gold-standard
\item quantified posterior probability contained within approximate intervals for $\logit^{-1}(h(\log \theta_i))$
\item quantified number of posterior evaluations
\item quantified computing time
\item use ``plasmode'' simulation to assess quality of method across variety of individuals (PK heterogeneity in this pop)
\end{itemize}
\section{Appendix: Prior hyperparameters for piperacillin among critically ill}
\begin{center}
\begin{table}
\begin{tabular}{lll} \hline
Parameter & Value \\ \hline
$\mu_0$ & \\
$\Sigma_0$ & \\
$\alpha_0$ & \\
$\beta_0$  & \\
\hline
\end{tabular}
\caption{Prior hyperparameters for piperacilling in a population of hospitalized, critically ill patients.}\label{tab:hyp}
\end{table}
\end{center}

\end{document}
